---
title: "Keegan Ds2020 Final"
author: "Keegan Moerke"
date: "2024-12-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Abstract

The goal of this project is to find trends in the walkability of cities in Iowa. Walkability is a key indicator of how well cities are planned and of the quality of life in that area. I chose to limit my project to just Iowa because the initial dataset was simply too large and Iowa is of relevance to all of us. Iowa also has plenty of rural and urban areas to show the difference.

```{r}
library(ggplot2)

```

This is the fullset, it is too big for github and to analyze, so I restrict it to only include Iowa and write it to another csv later. Here is the source to the data.
https://catalog.data.gov/dataset/walkability-index7
```{r}

df <-  read.csv("EPA_SmartLocationDatabase_V3_Jan_2021_Final.csv")

head(df)
```



Below shows restricting the data to just Iowa, then several plots comparing variables to walkability.
```{r}
 # This will install ggplot2 and other useful packages
library(tidyverse)

iowa_df = subset(df, `STATEFP` == 19)
# Assuming df is your dataframe
ggplot(data = iowa_df, aes(x = TotPop, y = NatWalkInd)) +
  geom_point() +
  labs(title = "Comparison of Total Population and National Walkability Index",
       x = "Total Population",
       y = "National Walkability Index") +
  theme_minimal()

ggplot(data = iowa_df, aes(x = AutoOwn0, y = NatWalkInd)) +
  geom_point() +
  labs(title = "Comparison of number of households owning zero vehicles and National Walkability Index",
       x = "Num Households owning zero vehicles",
       y = "National Walkability Index") +
  theme_minimal()




# Convert data from wide to long format
long_df <- iowa_df %>%
  pivot_longer(cols = c(E_LowWageWk, E_MedWageWk, E_HiWageWk), names_to = "Variable", values_to = "Value")

# Plot the reshaped data
ggplot(data = long_df, aes(x = Value, y = NatWalkInd, color = Variable)) +
  geom_point() +
  labs(title = "Comparison of Walkability Index with Workers of different income",
       x = "Workers separated by wage classification",
       y = "National Walkability Index",
       color = "Variable") +
  theme_minimal()



```
The first plot just shows population of a block to its walkability index. There is not really any correlation.

The second plot compares the number of households that own zero cars to the walkability index. This shows that there is some correlation. In the block with the most households without cars have fairly high walkability.

The third plot attempts to find if there is a correlation between walkability and the income of its residents. This shows that the more walkable blocks have higher income residents.

```{r}
# Writing the dataframe to a CSV file
write.csv(iowa_df, file = "iowa_walkability_data.csv", row.names = FALSE)

```
```{r}

# Assuming 'iowa_df' is your dataframe and it has columns 'COUNTYFP' and 'NatWalkInd'
ggplot(data = iowa_df, aes(x = factor(CSA_Name), y = NatWalkInd)) +
  geom_boxplot() +
  labs(title = "Faceted Boxplot of Walkability by County",
       x = "County",
       y = "National Walkability Index") +
  facet_wrap(~ CSA_Name, nrow = 4) +
  theme_minimal()

```

This plot shows the distribution of walkability within a certain census-designated area such as the Des Moines area. This shows that places like Des Moines has a higher walkability than places that are not associated with a metropolitan area.

```{r}
library(ggplot2)
ggplot(data = iowa_df, aes(x = D3B, y = NatWalkInd)) +
  geom_point() +
  labs(title = "Comparison of Intersection Density and National Walkability Index",
       x = "Intersection Density",
       y = "National Walkability Index") +
  theme_minimal()
```
This plot compares intersection density to walkability score


```{r}
library(ggplot2)
# Histogram of National Walkability Index
ggplot(data = iowa_df, aes(x = NatWalkInd, weight = TotPop)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of National Walkability Index weighted by population",
       x = "National Walkability Index",
       y = "Frequency") +
  theme_minimal()

```
This plot shows distribution of each block by its walkability indexs weighted by population


```{r}
library(corrplot)
relevant <- iowa_df %>% select(TotPop, NatWalkInd, D3B, D3BPO3, Workers, AutoOwn0, E5_Ret, E5_Off, E5_Ind, D4D, D4E, D5BE )

numeric_data <- iowa_df %>% select_if(is.numeric)

# Compute correlations with NatWalkInd
correlations <- sapply(numeric_data, function(x) cor(x, iowa_df$NatWalkInd, use = "complete.obs"))

# Convert to dataframe for better display
cor_df <- data.frame(Variable = names(correlations), Correlation = correlations)

# Display the correlations sorted by absolute value
cor_df <- cor_df %>% arrange(desc(abs(Correlation)))

print(cor_df)



```


## Conclusion

Walkability depends on multiple factors, some of which can be determined by city planners and developers. When developing  new neighborhoods, the should consider sticking to a simple grid pattern and keep the lots a modest size. This will increase the intersection density and thus the walkability. Additionally, cities should consider adding more transit options such as buses to give residents more options other than driving a personal vehicle. The walkability data also tends to show that wealthy areas have better walkability. Poor walkability puts everyone at a disadvantage.