---
title: "DS 202 Final Project"
output: 
  github_document:
    number_section: FALSE
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = TRUE)


knit_hooks$set( fig.cap = function(before, options, envir) {
  if ( before ) {
    chr <- sprintf("<p><small><strong><a name='fig:%s' value = %d>Figure %d</a></strong>: %s</small></p>", options$label, options$fig.num, options$fig.num, options$fig.cap)
    chr <- sprintf("<p><small><strong><a name='fig:%s'>%s</a></strong>: %s</small></p>", options$label, options$label,  options$fig.cap)
    
    return(chr)
  }
})


chunkref <- local({
  function(chunklabel) {
    sprintf('[%s](#%s)', chunklabel, paste0("fig:",chunklabel) )
  }  
})

library(ggplot2)
library(dplyr)
```


This repository serves as a starter repo for your final project, and this Rmd is supposed to serve as a starter file for your project report.

## Part I: Repo Structure {.unnumbered}

The structure sketched out below is an idea of what your repository might look like. You can use it as a starting base and change according to your needs. But think about the changes that you make!

    -- code
    |   |   -- any R scripts you need but don't want to include directly in the write-up
    -- data
    |   |   -- csv files (cleaned data)
    -- data-raw
    |   |   -- raw data files 
    |   |   -- data description files, origin
    |   |   -- Codebook
    -- final-project.Rmd
    -- images  # only images that are not created by the Rmd
    -- LICENSE
    -- README.md
    -- README.Rmd
    -- README_files # folder with files created during the knitting process

## Part II: Project report {.unnumbered}



# Title of your project {.unnumbered}

Authors: First Name Last Name, First Name Last Name, ...

## Abstract (TL;DR) {.unnumbered}

An abstract is a quick summary of your work. Ideally it should motivate someone to read the rest of the paper. Include one sentence each on

-   what is the project about?
-   what is the motivation for doing it?
-   what data is your work based on? and where does it come from? = what are your main findings? (one sentence each)

# Intro/Background/Motivation

What is the topic of your project, why is it relevant?

At the end of the Intro, write a sentence describing what each of the (result) sections is about, e.g. in section [Results 1] we show the relationship between XXX and YYY,  section [Results 2] also considers the effect of variable ZZZ. ...
Finally we conclude with a quick summary of our findings and potential follow-up work in section [Conclusions].

Somewhere at the beginning of your project, include a code chunk that includes all of the R packages you are using throughout. In this document, the setup code chunk is called `setup` (see line 8) Also make sure to set defaults for the code chunks - like should they be visible? (probably not: echo=FALSE). Do you want to automatically include warnings? (probably yes, for creating the Rmd, to make sure that all warnings are accounted for)

# Quick Data Summary

What are the variables that you will be using in the main part of the report? What are their ranges? You could include a table with variable names, a short explanation, and (very broad) summary statistics.

# Results

Each line of exploration is supposed to be featured in one of the Results sections. Make sure to change to more interesting section headers!

## Results 1

In your write-up, make sure to refer to all of the figures you create. You can include a hyperlink to the [scatterplot](#fig:scatterplot) by using the name of the code chunk (make sure, to give each code chunk a different name). In your markdown document you can create this link either by calling the function `chunkref` with the name of the code chunk in quotes, i.e. `` r chunkref("scatterplot") `` or by using the markdown expression `[scatterplot](#fig:scatterplot)`. Similarly, we can refer to the `r chunkref("2nd scatterplot")`. Note that the figure captions appear above the figures - this saves us from having to scroll up after following the link.

```{r scatterplot, echo=FALSE, fig.cap="This is the figure caption. Make sure to use the description we practised in the homework: first sentence describes structure of the plot, second sentence describes main finding, third sentence describes outliers/follow-up.", message = FALSE}

mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) + 
  geom_point()
```

```{r 2nd scatterplot, echo=FALSE, fig.cap="This is the figure caption. Make sure to use the description we practised in the homework: first sentence describes structure of the plot, second sentence describes main finding, third sentence describes outliers/follow-up.", message = FALSE}

mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) + 
  geom_point()
```

Additionally, you can also refer to different sections in your writeup by using anchors (links) to section headers. Here, we are referring to subsection [Results 3]. The code for that is `[Results 3]`.

## Results 2

## Results 3

...

# Conclusions

Give a quick summary of your work. Here is the place to be a bit critical and discuss potential limitations. Add a sentence on what else you would have liked to include in your data exploration if you had more time or more members in your team. 

## Data source {.unnumbered}

Where does the data come from, who owns the data? Where are all the scripts that you need to clean the data?

## References {.unnumbered}

List all resources you used.
